Fixing Vulnerabilities Hinders Maintainability

### 2019-02-04 paper/ICPC19/icpc19.tex
review partial

\newcommand\Sof[1]{\nb{Sofia}{red}{#1}}
\newcommand\Luis[1]{\nb{Luis}{green}{#1}}
\newcommand\Rui[1]{\nb{Rui}{blue}{#1}}

@@

\end{lstlisting} \Luis{no listing lst:fix o que achas de fazer linebreak para
cada uma das tags? assim nao havi o problema de ter a mesma linha dividida e ficava mais facil de ler:\\
\texttt{<p class='hint'>}\\
\texttt{~~<?php }\\
\texttt{</p>}\\
Also: reduz numero de espacos para cada tab -- 2 é suficiente.
}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The
methodology is comprised of the following steps, as illustrated in
Figure~\ref{fig:met}:

@@

    \item \textbf{Miscellaneous (195 cases)} This pattern comprises several other categories
    of security refactorings (e.g., \textit{path traversal and buffer
    overflow}). It contains several other security refactorings that do not have
    their own pattern yet. The patterns not satisfying the size requirement of
    more than $20$ refactorings are the following ones: \textit{Resource Leak} (19),
    \textit{SHA-1 Hash Function} (1), \textit{Broken Access Control} (2),
    \textit{Sensitive Data Exposure} (18), \textit{Insufficient Attack Protection} (4),
    \textit{Overflow} (18), \textit{Path Traversal} (19).
\end{itemize}

@@

The dataset contains security flaws for more than $13$ different languages,
being PHP ($39\%$) and C ($24\%$) the most prevalent ones (see Figure~\ref{fig:lang}).
To classify the applications in the dataset, we use a taxonomy for open-source
software from previous work~\cite{7816479}. Figure~\ref{fig:domain}
presents the projects domain distribution: \textit{Aplication Software} ($25$),
software that provides end-users with functional systems; \textit{Web libraries
and frameworks} ($20$); \textit{System Software} ($18$), software that provides
services and infrastructures (e.g., operating systems, servers and databases);
\textit{Software Tool} ($13$), software that supports development (e.g.,
programming languages, compilers, package managers, IDEs); and, \textit{Non-web
libraries and frameworks} ($18$), for desktop and mobile software.
%
Table~\ref{tab:dataset} presents the descriptive statistics of the $94$
open-source projects involved in this study, including the number of forks, the
number of stars, the number of watchers, the number of contributors, the number
of commits, the number of branches, the number of releases,  size (cf. GitHub, it
is the size of the whole repository, including all of its history, in
kilobytes), the number of issues, and the number of pull requests.

>>>>

\newcommand\Sof[1]{\nb{Sofia}{red}{#1}}
\newcommand\Luis[1]{\nb{Luis}{mygreen}{#1}}
\newcommand\Rui[1]{\nb{Rui}{blue}{#1}}

@@

\end{lstlisting} \Luis{no listing lst:fix o que achas de fazer linebreak para
cada uma das tags? assim nao havi o problema de ter a mesma linha dividida e ficava mais facil de ler:\\
\texttt{<p class='hint'>}\\
\texttt{~~<?php }\\
\texttt{</p>}\\
Also: reduz numero de espacos para cada tab -- 2 é suficiente.
}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The
methodology is comprised of the following steps, as illustrated in
Figure~\ref{fig:met}\Luis{como usamos numeros na enumeracao que descreve a
figura, esses numeros tb deveriam estar na figura.}:

@@

    \item \textbf{Miscellaneous (195 cases)} This pattern comprises several
    other categories of security refactorings (e.g., \textit{path traversal and
    buffer overflow})\Luis{exemplos ja nao precisa estar aqui porque sao
    enumerados logo a seguir}. It contains several other security refactorings
    that do not have their own pattern yet. The patterns not satisfying the
    size requirement of more than $20$ refactorings are the following ones:
    \textit{Resource Leak} (19), \textit{SHA-1 Hash Function} (1),
    \textit{Broken Access Control} (2), \textit{Sensitive Data Exposure} (18),
    \textit{Insufficient Attack Protection} (4), \textit{Overflow} (18),
    \textit{Path Traversal} (19).\Luis{Acho que este item pode ser reescrito de forma mais clara. sugestao:
    This category comprises the security refactorings that (1) do not relate to
    any particular pattern or (2) relate to patterns that do not occur in at
    least 20 refactorings: \textit{Resource Leak} (19), \dots} \end{itemize}

@@

The dataset contains security flaws for more than $13$ different languages,
being PHP ($39\%$) and C ($24\%$) the most prevalent ones (see Figure~\ref{fig:lang}).
To classify the applications in the dataset, we use a taxonomy for open-source
software from previous work~\cite{7816479}. Figure~\ref{fig:domain}
presents the projects domain distribution: \textit{Application Software} ($25$),
software that provides end-users with functional systems; \textit{Web libraries
and frameworks} ($20$); \textit{System Software} ($18$), software that provides
services and infrastructures (e.g., operating systems, servers and databases);
\textit{Software Tool} ($13$), software that supports development (e.g.,
programming languages, compilers, package managers, IDEs); and, \textit{Non-web
libraries and frameworks} ($18$), for desktop and mobile software.
%
Table~\ref{tab:dataset} presents the descriptive statistics of the $94$
open-source projects involved in this study, including the number of forks, the
number of stars, the number of watchers, the number of contributors, the number
of commits, the number of branches, the number of releases,  size (cf. GitHub, it
is the size of the whole repository, including all of its history, in
kilobytes), the number of issues, and the number of pull requests.
